Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-11 arxiv/proof_same_cov.tex
sec A

	\paragraph{Proof of Claim \ref{claim_opt_dist}.}
	Corresponding to the right-hand side of \eqref{eq_gA}, we define the function
	$$h(A):= n \bignormFro{\Sigma^{1/2} B^{\star} \bigbrace{A^{\top} (AA^{\top})^{\dagger} A - \id_{t\times t}}}^2 + \sigma^2 (n\cdot t - p \cdot r)$$
	In order to show that the $U_{\hat{A}}U_{\hat{A}}^\top$ is close to $A^\star {A^\star}^\top$, we first show that $g(A)$ is close to $h(A)$ as follows:
	\begin{align}\label{eq_gA_err}
		\bigabs{g(A) - h(A)} \lesssim  n^{-c_{\varphi}} \cdot n \bignorm{\Sigma^{1/2} B^{\star} (U_AU_A^{\top} - \id_{t\times t})}^2 + n^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t,
	\end{align}
	for any fixed values $c_{\varphi}$ within $(0, \frac{\varphi-4}{2\varphi})$ and $c_{\infty}$ within $(0, 1/2)$. 
%	\begin{align}\label{eq_gA_err}
%		\bigabs{g(A) - \exarg{\cE, X}{g(A)}} \lesssim p^{-c_{\varphi}} \cdot n \bignorm{\Sigma^{1/2} B^{\star} (U_AU_A^{\top} - \id)}^2 + p^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t.
%	\end{align}
	To show the above result, we consider the concentration error of each part of $g(A)$.

@@

	\bigskip
	Next, we use equation \eqref{eq_gA_err} to prove the claim.
	Using triangle inequality, we upper bound the gap between $g(A^{\star})$ and $g(\hat{A})$:
	\begin{align}
		g(A^{\star}) - g(\hat{A}) &\le \bigabs{g(A^{\star}) - h(A^{\star})} + (h(A^{\star}) - h(\hat{A})) + \big|g(\hat{A}) - h(\hat{A})\big| \nonumber \\
		&\le \bigabs{g(A^{\star}) - h(A^{\star})}  + \big|g(\hat{A}) - h(\hat{A})\big|\nonumber \\
		&\lesssim n^{-c_\varphi}\cdot n \bignormFro{\Sigma^{1/2} B^{\star}}^2 + n^{-c_\infty}\cdot \sigma^2 \cdot n \cdot t. \label{eq_g_gap}
	\end{align}
	The second step used the fact that $A^{\star}$ is a global minimum of $h(\cdot)$, so that $h(A^{\star}) \le h(\hat{A})$.
	The third step used equation \eqref{eq_gA_err} and the fact that the spectral norm of $U_A U_A^{\top} - \id$ is at most one.

	Using \eqref{eq_g_gap}, we can derive an upper bound on $\bignormFro{A^{\star} {A^{\star}}^{\top}- U_{\hat{A}} U_{\hat{A}}^{\top}}$ as follows.
	We simplify equation \eqref{eq_mtl_output_layer} as
	\begin{align*}
		g(A)  &= \normFro{X (X^{\top}X)^{-1} X^{\top} Y U_A U_A^{\top} - Y}^2 \\
					&= \normFro{Y}^2 - \inner{U_A U_A^{\top}}{Y^{\top} X (X^{\top} X)^{-1} X^{\top} Y}.
	\end{align*}
	We denote $H = Y^{\top} X (X^{\top} X)^{-1} X^{\top} Y$, which appears in the above equation.
	Since $\hat{A}$ is a global minimum of $g(A)$, we have that $g(A^{\star}) \ge g(\hat{A})$.
	Hence, we have
	\begin{align}
		g(A^{\star}) - g(\hat{A}) &= \abs{g(A^{\star}) - g(\hat{A})} = \abs{\inner{H}{A^{\star} {A^{\star}}^{\top} - U_{\hat{A}} U_{\hat{A}}^{\top}}} \nonumber \\
												&\ge \lambda_{\min}(H) \cdot \normFro{A^{\star} {A^{\star}}^{\top} - U_{\hat{A}} U_{\hat{A}}^{\top}}. \label{eq_g_gap1}
	\end{align}
	Recall that $Y = X B^{\star} + \cE$ in matrix notation.
	We bound the minimum singular value of $H$ as follows
	\begin{align}
		\lambda_{\min}(H) &= \lambda_{\min}({B^{\star}}^{\top} X^{\top} X B^{\star} + \cE^{\top} X (X^{\top} X)^{-1} X^{\top} \cE) \nonumber \\
		&\ge \lambda_{\min}({B^{\star}}^{\top}X^{\top}X B^{\star}) + \lambda_{\min}(\cE^{\top} X (X^{\top} X)^{-1} X^{\top} \cE) \nonumber \\
		&\ge \lambda_{\min}({B^{\star}}^{\top}\Sigma B^{\star}) \cdot \lambda_{\min}(Z^{\top}Z) + \lambda_{\min}(\cE^{\top} X (X^{\top} X)^{-1} X^{\top} \cE) \nonumber \\
		&\ge \lambda_{\min}({B^{\star}}^{\top}\Sigma B^{\star}) \cdot ((\sqrt n - \sqrt p)^2 - n \cdot p^{-c}) + \sigma^2 \cdot p \cdot (1 - p^{-1/2 + \e}). \label{eq_g_gap2}
	\end{align}
	The second equation uses the fact that the minimum singular value of the sum of two PSD matrices is greater than the sum of the minimum singular value of each matrix.
	The third equation uses the fact that $X = Z \Sigma^{1/2}$ and Fact \ref{fact_proof_gA} in Appendix \ref{app_tool}.
	The last equation uses Fact \ref{fact_minv} (ii) for the minimum singular value of $Z^{\top}Z$ and Lemma \ref{largedeviation} for $\cE\cE^{\top}$.

	Combining equation \eqref{eq_g_gap}, \eqref{eq_g_gap1}, and \eqref{eq_g_gap2}, we conclude that
	\begin{align*}
		\bignormFro{A^{\star} {A^{\star}}^{\top}- U_{\hat{A}} U_{\hat{A}}^{\top}}
		&\le \frac{p^{-c} \cdot n \bignormFro{\Sigma^{1/2} B^{\star}}^2 + p^{-1/2 + c} \cdot \sigma^2 nt} {\lambda_{\min}(H)} \\
		&\le \frac{p^{-c} \rho \normFro{\Sigma^{1/2} B^{\star}}^2 + p^{-1/2 + \e} \cdot \sigma^2 \rho t}{\lambda_{\min}^2(\Sigma^{1/2} B^{\star}) ((\sqrt{\rho} - 1)^2 - \rho \cdot p^{-c}) + \sigma^2 (1 - p^{-1/2 + \e})}
	\end{align*}
	Hence for large enough $p$, we obtain that Claim \ref{claim_opt_dist} holds.
	The proof is complete.

>>>>

	\paragraph{Proof of Claim \ref{claim_opt_dist}.}
	Corresponding to the right-hand side of \eqref{eq_gA}, we define the function
	\be\label{same_hA}h(A):= n \bignormFro{\Sigma^{1/2} B^{\star} \bigbrace{A^{\top} (AA^{\top})^{\dagger} A - \id_{t\times t}}}^2 + \sigma^2 (n\cdot t - p \cdot r).\ee
	In order to show that the $U_{\hat{A}}U_{\hat{A}}^\top$ is close to $A^\star {A^\star}^\top$, we first show that $g(A)$ is close to $h(A)$ as follows:
	\begin{align}\label{eq_gA_err}
		\bigabs{g(A) - h(A)} \lesssim  n^{-c_{\varphi}} \cdot n \bignorm{\Sigma^{1/2} B^{\star} (U_AU_A^{\top} - \id_{t\times t})}^2 + n^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t,
	\end{align}
	for any fixed values $c_{\varphi}$ within $(0, \frac{\varphi-4}{2\varphi})$ and $c_{\infty}$ within $(0, 1/2)$. 
%	\begin{align}\label{eq_gA_err}
%		\bigabs{g(A) - \exarg{\cE, X}{g(A)}} \lesssim p^{-c_{\varphi}} \cdot n \bignorm{\Sigma^{1/2} B^{\star} (U_AU_A^{\top} - \id)}^2 + p^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t.
%	\end{align}
	To show the above result, we consider the concentration error of each part of $g(A)$.

@@

	\bigskip
	Next, we use equation \eqref{eq_gA_err} to prove the claim.
	Using triangle inequality, we upper bound the gap between $g(A^{\star})$ and $g(\hat{A})$:
	\begin{align}
		h(\hat{A})- h(A^{\star})   &\le \bigabs{g(A^{\star}) - h(A^{\star})} + (g(\hat A) - g({A}^{\star})) + \big|g(\hat{A}) - h(\hat{A})\big| \nonumber \\
		&\le \bigabs{g(A^{\star}) - h(A^{\star})}  + \big|g(\hat{A}) - h(\hat{A})\big|\nonumber \\
		&\lesssim n^{-c_\varphi}\cdot n \bignormFro{\Sigma^{1/2} B^{\star}}^2 + n^{-c_\infty}\cdot \sigma^2 \cdot n \cdot t. \label{eq_g_gap}
	\end{align}
	The second step used the fact that $\hat A$ is the global minimizer of $g(\cdot)$, so that $g(\hat A) \le g({A}^\star)$.
	The third step used equation \eqref{eq_gA_err} and the fact that the spectral norm of $U_A U_A^{\top} - \id$ is at most one. 
	%With \eqref{eq_g_gap}, we can derive an upper bound on $\bignormFro{A^{\star} {A^{\star}}^{\top}- U_{\hat{A}} U_{\hat{A}}^{\top}}$ as follows. 
	Using equation \eqref{same_hA}, we can obtain that
	$$h(\hat A)-h(A^\star) = n \bigtr{{B^\star}^\top\Sigma B^\star ( A^\star {A^\star}^\top -U_{\hat A}U^\top_{\hat A})} .$$
	Let $\lambda_1\ge\lambda_2 \ge \cdots\ge \lambda_t$ be the eigenvalues of ${B^\star}^\top\Sigma B^\star$, and $v_i$, $1\le i \le t$, be the corresponding eigenvectors. Then we have $A^\star {A^\star}^\top =\sum_{i=1}^r v_i v_i^\top$, and 
	\begin{align}
	h(\hat A)-h(A^\star) = n \sum_{i=1}^r \lambda_i - n\sum_{i=1}^t \lambda_i \| U^\top_{\hat A} v_i\|^2 &= n\sum_{i=1}^r \lambda_i\left(1 -  \| U^\top_{\hat A} v_i\|^2\right)-n\sum_{i=r+1}^t \lambda_i \| U^\top_{\hat A} v_i\|^2 \nonumber\\
	&\ge  n(\lambda_r-\lambda_{r+1}) \sum_{i=r+1}^t \| U^\top_{\hat A} v_i\|^2 , \label{bdd_A-A}
	\end{align}
	where we used $\sum_{i=1}^r \left(1 -  \| U^\top_{\hat A} v_i\|^2\right) = r- \sum_{i=1}^r  \| U^\top_{\hat A} v_i\|^2 =\sum_{i=r+1}^t \| U^\top_{\hat A} v_i\|^2  $ in the last step. On the other hand, we have 
	\begin{align*}
	\| A^\star {A^\star}^\top -U_{\hat A}U^\top_{\hat A}\|_F^2 & =\sum_{i=1}^t v_i^\top (A^\star {A^\star}^\top -U_{\hat A}U^\top_{\hat A})^2 v_i \\
	& =\sum_{i=1}^r \left(1 -  \| U^\top_{\hat A} v_i\|^2\right) + \sum_{i=r+1}^t \| U^\top_{\hat A} v_i\|^2 =2 \sum_{i=r+1}^t \| U^\top_{\hat A} v_i\|^2.
	\end{align*}
	Thus from \eqref{bdd_A-A}, we obtain that 
	$$\| A^\star {A^\star}^\top -U_{\hat A}U^\top_{\hat A}\|_F^2  =2 \sum_{i=r+1}^t \| U^\top_{\hat A} v_i\|^2 \lesssim \frac{n^{-c_\varphi}\cdot  \bignormFro{\Sigma^{1/2} B^{\star}}^2 + n^{-c_\infty}\cdot \sigma^2 t}{\lambda_r- \lambda_{r+1}}.$$
	This completes the proof of Claim \ref{claim_opt_dist}.
	\iffalse
	We simplify equation \eqref{eq_mtl_output_layer} as
	\begin{align*}
		g(A)  &= \normFro{X (X^{\top}X)^{-1} X^{\top} Y U_A U_A^{\top} - Y}^2 \\
					&= \normFro{Y}^2 - \inner{U_A U_A^{\top}}{Y^{\top} X (X^{\top} X)^{-1} X^{\top} Y}.
	\end{align*}
	We denote $H = Y^{\top} X (X^{\top} X)^{-1} X^{\top} Y$, which appears in the above equation.
	Since $\hat{A}$ is a global minimum of $g(A)$, we have that $g(A^{\star}) \ge g(\hat{A})$.
	Hence, we have
	\begin{align}
		g(A^{\star}) - g(\hat{A}) &= \abs{g(A^{\star}) - g(\hat{A})} = \abs{\inner{H}{A^{\star} {A^{\star}}^{\top} - U_{\hat{A}} U_{\hat{A}}^{\top}}} \nonumber \\
												&\ge \lambda_{\min}(H) \cdot \normFro{A^{\star} {A^{\star}}^{\top} - U_{\hat{A}} U_{\hat{A}}^{\top}}. \label{eq_g_gap1}
	\end{align}
	Recall that $Y = X B^{\star} + \cE$ in matrix notation.
	We bound the minimum singular value of $H$ as follows
	\begin{align}
		\lambda_{\min}(H) &= \lambda_{\min}({B^{\star}}^{\top} X^{\top} X B^{\star} + \cE^{\top} X (X^{\top} X)^{-1} X^{\top} \cE) \nonumber \\
		&\ge \lambda_{\min}({B^{\star}}^{\top}X^{\top}X B^{\star}) + \lambda_{\min}(\cE^{\top} X (X^{\top} X)^{-1} X^{\top} \cE) \nonumber \\
		&\ge \lambda_{\min}({B^{\star}}^{\top}\Sigma B^{\star}) \cdot \lambda_{\min}(Z^{\top}Z) + \lambda_{\min}(\cE^{\top} X (X^{\top} X)^{-1} X^{\top} \cE) \nonumber \\
		&\ge \lambda_{\min}({B^{\star}}^{\top}\Sigma B^{\star}) \cdot ((\sqrt n - \sqrt p)^2 - n \cdot p^{-c}) + \sigma^2 \cdot p \cdot (1 - p^{-1/2 + \e}). \label{eq_g_gap2}
	\end{align}
	The second equation uses the fact that the minimum singular value of the sum of two PSD matrices is greater than the sum of the minimum singular value of each matrix.
	The third equation uses the fact that $X = Z \Sigma^{1/2}$ and Fact \ref{fact_proof_gA} in Appendix \ref{app_tool}.
	The last equation uses Fact \ref{fact_minv} (ii) for the minimum singular value of $Z^{\top}Z$ and Lemma \ref{largedeviation} for $\cE\cE^{\top}$.

	Combining equation \eqref{eq_g_gap}, \eqref{eq_g_gap1}, and \eqref{eq_g_gap2}, we conclude that
	\begin{align*}
		\bignormFro{A^{\star} {A^{\star}}^{\top}- U_{\hat{A}} U_{\hat{A}}^{\top}}
		&\le \frac{p^{-c} \cdot n \bignormFro{\Sigma^{1/2} B^{\star}}^2 + p^{-1/2 + c} \cdot \sigma^2 nt} {\lambda_{\min}(H)} \\
		&\le \frac{p^{-c} \rho \normFro{\Sigma^{1/2} B^{\star}}^2 + p^{-1/2 + \e} \cdot \sigma^2 \rho t}{\lambda_{\min}^2(\Sigma^{1/2} B^{\star}) ((\sqrt{\rho} - 1)^2 - \rho \cdot p^{-c}) + \sigma^2 (1 - p^{-1/2 + \e})}
	\end{align*}
	Hence for large enough $p$, we obtain that Claim \ref{claim_opt_dist} holds.
	\fi
